Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2019-02-07 WholePlotTrajectories.Rmd
intro

The large areas covered with tropical forests worldwide hold crucial environmental, economic and social values. 
They provide wood and multiple non-timber forest products, shelter a diversified fauna, regulate the local and regional climates, the carbon, water and nutrient cycles, and ensure cultural and human well-being. 
The growing demand in forests products together with current global changes, however, increase the pressure on remaining natural forests [@Morales-Hidalgo2015].
This changing context alters the natural disturbance regime that shape and maintain tree community structure, composition and functioning [@Schnitzer2001;@Anderson-Teixeira2013;@Sist2015].
Anticipate the fate of tropical forests then requires to understand the determinants of community response to disturbance and the underlying processes.
In tropical forests, disturbance is followed by secondary succession that starts with a maintained forest environment.
Post-disturbance successional trajectories follow environmental variability and comprises a suit of changes in community structure and functioning.
This succession is driven by various processes involving the different species responses to disturbance (as colonisation, facilitation, coexistence, etc) [@Lindenmayer2012;@Garcia_florez2017].
The rate, endpoint an duration of these trajectories depend on the intensity of the disturbance: a current aim of tropical ecology is to highlight the linkage between biodiversity and disturbance and infer predictable trajectories [@Chazdon2003a;@Willig2018].

Disturbance change both community abiotic and biotic environments, through modifications in the fluxes of light, heat and water [@Goulamoussene2017]. 
For now, post-disturbance trajectories has been largely studied through forest structural parameters such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016]. 
Models based on observed trajectories then assessed community time to recovery and highlighted the role of forest structure, forest composition, environmental parameters and disturbance characteristics for the determination of post-disturbance trajectories [@Herault2018]. 
Regarding community diversity and composition, however, post-disturbance trajectories have not been as thoroughly understood [@Guitet2018;@Molino2001].
There is a need to further analyse community post-disturbance trajectories to assess community recovery and linkage between biodiversity and disturbance.

An early conceptual basis of the linkage between biodiversity and disturbance is the Intermediate Disturbance Hypothesis (IDH).
The IDH assumes a relationship between community diversity and the intensity and frequency of disturbance events [@Connell1978], and states that community diversity peaks at intermediate level of disturbance.
Post-disturbance fluctuations of environmental conditions are assumed to prevent competitive exclusion in fostering both competitively superior species and fast colonizers [@Shea2004;@Pulsford2016].
In highly diverse tropical forests, however, the validation of the IDH remains controversial [@Hubbell2001;@Fox2013;@Sheil2013] and observations are often blurry compared to theoretical expectations [@Hugues2007;@Sheil2003;@Norden2017].
The IDH pattern of peaked diversity under intermediate disturbance can result from various processes that should be identified to grasp the conditions for the IDH application and the impact on communities, and to anticipate disturbance outcomes [@Sheil2003;@Shea2004].

>>>>

The large areas covered with tropical forests worldwide hold crucial environmental, economic and social values. 
They provide wood and multiple non-timber forest products, shelter a diversified fauna, and ensure cultural and human well-being.
They regulate as well the local and regional climates, and the carbon, water and nutrient cycles.
However, the growing demand in forests products together with current global changes increase the pressure on remaining natural forests [@Morales-Hidalgo2015].
This growing pressure affects the natural disturbance regime that shape and maintain the structure, composition and functioning of tree communities [@Schnitzer2001;@Anderson-Teixeira2013;@Sist2015].
It is then urgent to understand the processes underlying community response to disturbance in order to anticipate the fate of tropical forests in the current context.
In tropical forests, disturbance maintains a forest environment and is followed by secondary succession comprising a suit of changes in the structure and functioning of communities.
Disturbance change both community abiotic and biotic environments, through modifications in the fluxes of light, heat and water [@Goulamoussene2017]. 
For now, post-disturbance trajectories has been largely studied through forest structural parameters such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016]. 
Models based on observed trajectories then assessed community time to recovery and highlighted the determinants of post-disturbance trajectories, as forest structure, forest composition, environmental parameters and disturbance characteristics [@Herault2018]. 
Regarding community diversity and composition, however, post-disturbance trajectories have not been as thoroughly understood [@Guitet2018;@Molino2001].
Biodiversity trajectories result from the differences in species response to disturbance which entail a variety of processes like colonisation, facilitation, coexistence [@Lindenmayer2012;@Garcia_florez2017].
The rate, endpoint and duration of post-disturbance trajectories depend on the intensity and frequency of disturbance: a current objective of tropical ecology is to highlight this linkage betwenn biodiversity and disturbance, and anticipate community response [@Chazdon2003a;@Willig2018].

An early conceptual basis of the linkage between biodiversity and disturbance is the Intermediate Disturbance Hypothesis (IDH).
The IDH assumes a relationship between community diversity and the intensity and frequency of disturbance events [@Connell1978], and states that community diversity peaks at intermediate level of disturbance.
The fluctuations of environmental conditions after disturbance are assumed to prevent competitive exclusion as they foster both competitively superior species and fast colonizers [@Shea2004;@Pulsford2016].
In highly diverse tropical forests, however, observations of the IDH are often blurry compared to theoretical expectations [@Hugues2007;@Sheil2003;@Norden2017] the validation of the theory remains controversial [@Hubbell2001;@Fox2013;@Sheil2013].
As previously highlighted, there is a variety of ecological processes emerging after disturbance, and the pattern of the IDH (the diversity peak under intermediate disturbance) can result from various processes.
Disentangle these processes in the context of tropical forests would assess the conditions for the IDH application and clarify the disturbance impact on communit diversity [@Sheil2003;@Shea2004].

### 2019-02-07 WholePlotTrajectories.tex
intro

The large areas covered with tropical forests worldwide hold crucial
environmental, economic and social values. They provide wood and
multiple non-timber forest products, shelter a diversified fauna,
regulate the local and regional climates, the carbon, water and nutrient
cycles, and ensure cultural and human well-being. The growing demand in
forests products together with current global changes, however, increase
the pressure on remaining natural forests \citep{Morales-Hidalgo2015}.
This changing context alters the natural disturbance regime that shape
and maintain tree community structure, composition and functioning
\citep{Schnitzer2001, Anderson-Teixeira2013, Sist2015}. Anticipate the
fate of tropical forests then requires to understand the determinants of
community response to disturbance and the underlying processes. In
tropical forests, disturbance is followed by secondary succession that
starts with a maintained forest environment. Post-disturbance
successional trajectories follow environmental variability and comprises
a suit of changes in community structure and functioning. This
succession is driven by various processes involving the different
species responses to disturbance (as colonisation, facilitation,
coexistence, etc) \citep{Lindenmayer2012, Garcia_florez2017}. The rate,
endpoint an duration of these trajectories depend on the intensity of
the disturbance: a current aim of tropical ecology is to highlight the
linkage between biodiversity and disturbance and infer predictable
trajectories \citep{Chazdon2003a, Willig2018}.

Disturbance change both community abiotic and biotic environments,
through modifications in the fluxes of light, heat and water
\citep{Goulamoussene2017}. For now, post-disturbance trajectories has
been largely studied through forest structural parameters such as
aboveground biomass, tree height or stem density
\citep{Piponiot2016, Rutishauser2016}. Models based on observed
trajectories then assessed community time to recovery and highlighted
the role of forest structure, forest composition, environmental
parameters and disturbance characteristics for the determination of
post-disturbance trajectories \citep{Herault2018}. Regarding community
diversity and composition, however, post-disturbance trajectories have
not been as thoroughly understood \citep{Guitet2018, Molino2001}. There
is a need to further analyse community post-disturbance trajectories to
assess community recovery and linkage between biodiversity and
disturbance.

An early conceptual basis of the linkage between biodiversity and
disturbance is the Intermediate Disturbance Hypothesis (IDH). The IDH
assumes a relationship between community diversity and the intensity and
frequency of disturbance events \citep{Connell1978}, and states that
community diversity peaks at intermediate level of disturbance.
Post-disturbance fluctuations of environmental conditions are assumed to
prevent competitive exclusion in fostering both competitively superior
species and fast colonizers \citep{Shea2004, Pulsford2016}. In highly
diverse tropical forests, however, the validation of the IDH remains
controversial \citep{Hubbell2001, Fox2013, Sheil2013} and observations
are often blurry compared to theoretical expectations
\citep{Hugues2007, Sheil2003, Norden2017}. The IDH pattern of peaked
diversity under intermediate disturbance can result from various
processes that should be identified to grasp the conditions for the IDH
application and the impact on communities, and to anticipate disturbance
outcomes \citep{Sheil2003, Shea2004}.

>>>>

The large areas covered with tropical forests worldwide hold crucial
environmental, economic and social values. They provide wood and
multiple non-timber forest products, shelter a diversified fauna, and
ensure cultural and human well-being. They regulate as well the local
and regional climates, and the carbon, water and nutrient cycles.
However, the growing demand in forests products together with current
global changes increase the pressure on remaining natural forests
\citep{Morales-Hidalgo2015}. This growing pressure affects the natural
disturbance regime that shape and maintain the structure, composition
and functioning of tree communities
\citep{Schnitzer2001, Anderson-Teixeira2013, Sist2015}. It is then
urgent to understand the processes underlying community response to
disturbance in order to anticipate the fate of tropical forests in the
current context. In tropical forests, disturbance maintains a forest
environment and is followed by secondary succession comprising a suit of
changes in the structure and functioning of communities. Disturbance
change both community abiotic and biotic environments, through
modifications in the fluxes of light, heat and water
\citep{Goulamoussene2017}. For now, post-disturbance trajectories has
been largely studied through forest structural parameters such as
aboveground biomass, tree height or stem density
\citep{Piponiot2016, Rutishauser2016}. Models based on observed
trajectories then assessed community time to recovery and highlighted
the determinants of post-disturbance trajectories, as forest structure,
forest composition, environmental parameters and disturbance
characteristics \citep{Herault2018}. Regarding community diversity and
composition, however, post-disturbance trajectories have not been as
thoroughly understood \citep{Guitet2018, Molino2001}. Biodiversity
trajectories result from the differences in species response to
disturbance which entail a variety of processes like colonisation,
facilitation, coexistence \citep{Lindenmayer2012, Garcia_florez2017}.
The rate, endpoint and duration of post-disturbance trajectories depend
on the intensity and frequency of disturbance: a current objective of
tropical ecology is to highlight this linkage betwenn biodiversity and
disturbance, and anticipate community response
\citep{Chazdon2003a, Willig2018}.

An early conceptual basis of the linkage between biodiversity and
disturbance is the Intermediate Disturbance Hypothesis (IDH). The IDH
assumes a relationship between community diversity and the intensity and
frequency of disturbance events \citep{Connell1978}, and states that
community diversity peaks at intermediate level of disturbance. The
fluctuations of environmental conditions after disturbance are assumed
to prevent competitive exclusion as they foster both competitively
superior species and fast colonizers \citep{Shea2004, Pulsford2016}. In
highly diverse tropical forests, however, observations of the IDH are
often blurry compared to theoretical expectations
\citep{Hugues2007, Sheil2003, Norden2017} the validation of the theory
remains controversial \citep{Hubbell2001, Fox2013, Sheil2013}. As
previously highlighted, there is a variety of ecological processes
emerging after disturbance, and the pattern of the IDH (the diversity
peak under intermediate disturbance) can result from various processes.
Disentangle these processes in the context of tropical forests would
assess the conditions for the IDH application and clarify the
disturbance impact on communit diversity \citep{Sheil2003, Shea2004}.
